---
title: "Week 5 Imputation"
author: "Tom Fitzpatrick"
date: "2025-08-13"
output: html_document
---

## Data Import and Summary

We import the dataset and take a quick look at the structure and overall proportions by gender.


# Load core packages
```{r,echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(mice)
library(broom)
```
# Load and inspect data
```{r }

referendum <- read_csv("/Users/tomfitzpatrick/Data201/referendum_survey_lab_version.csv")

```
# Overall counts and proportions by gender
```{r }
gender_table <- referendum %>%
  count(gender, referendum) %>%
  group_by(gender) %>%
  mutate(prop_yes = n / sum(n) * 100)
gender_table
```
# Fit complete-case logistic regression
```{r}

model_complete <- glm(referendum ~ age + gender, data = referendum, family = binomial())
summary(model_complete)
```


# Perform multiple imputation
```{r}
imputed_data <- mice(referendum %>% select(age, gender, referendum), m = 5, seed = 123)
```
# Fit logistic regression models to imputed data
```{r}
model_imputed <- with(imputed_data, glm(referendum ~ age + gender, family = binomial()))
```
# Pool results
```{r}
pooled_results <- pool(model_imputed)
```
# Get model terms for filtering
```{r}
all_terms <- tidy(model_complete)$term
gender_terms <- grep("^gender", all_terms, value = TRUE)
```
# Tidy data for complete-case model
```{r}
complete_tidy <- tidy(model_complete) %>%
  filter(term == "age" | term %in% gender_terms) %>%
  mutate(analysis = "Complete Cases")
```
# Tidy data for pooled imputed model
```{r}
imputed_tidy <- tidy(pooled_results) %>%
  filter(term == "age" | term %in% gender_terms) %>%
  mutate(analysis = "Multiple Imputation")
```
# Combine and recode terms for plotting
```{r}
combined <- bind_rows(complete_tidy, imputed_tidy) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = recode(term,
                       age = "Age",
                       `genderFemale` = "Gender: Female",
                       `genderMale` = "Gender: Male",
                       `genderGenderqueer, neither exclusively male nor female` = "Gender: Queer/Nonbinary",
                       `genderOther (Please State)` = "Gender: Other",
                       `genderTransgender Female (Male to Female: MTF)` = "Gender: Trans Female (MTF)",
                       `genderTransgender Male (Female to Male; FTM)` = "Gender: Trans Male (FTM)"
  )) %>%
  filter(std.error < 10)
```
# Plot estimates with error bars
```{r}
ggplot(combined, aes(x = term, y = estimate, color = analysis)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error),
                position = position_dodge(width = 0.5), width = 0.2) +
  labs(title = "Comparison of Logistic Regression Estimates",
       y = "Estimate (log-odds)",
       x = "Predictor",
       color = "Analysis Method") +
  theme_minimal()
```
